---
title: "EDA"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup}
# Load libraries
require(data.table)
require(magrittr)
require(tidyverse)
require(tidytext)
```

```{r load data}
# Train data
train <- fread("../data/train.csv")

# Test data
test <- fread("../data/test.csv")
```

# Data Exploration
```{r}
dim(train)
dim(test)
# Test is much larger than train

names(train)
names(test)

# Preview of data
head(train, 20)

# Unique qids in train
length(unique(c(train$qid1, train$qid2)))

# How many times do different ids appear?
table(c(train$qid1, train$qid2)) %>% 
  as.data.table %>% 
  .[,.N, by = N]

# How many ids appear in both locations?
sum(train$qid1 %in% train$qid2)

```

An important question here is what makes two questions the same? Questions are the same if the *main idea* is the same, even if the phrasing changes. The algorithm needs to be able to pick up on the *main idea* of each question and determine how much they overlap.

## Possible features
The following is a list of features that may be helptful to incorportate:
* Shared words
* Sentiment of each question
* Shared words weighted by importance
* Shared subject
* word2vec features - what if all text is thrown into a copus and passed through word2vec? Numerical similarities could be computed for each question using the vector embeddings for each word - how far is the center of this sentence from the center of that sentence?

```{r tidytext}

```




